Generative Adversarial Networks

### 2017-12-04 cs6735/project/report.tex
small updates (computer is crashing)

\subsection{Image Quality}
When generating images, problems often arise due to the generators difficulties with perspective, counting and structure \citep{2017arXiv170100160G}. For example, an generated animal may  have a 2D structure, misplaced limbs or too many of a specific body part. These issues arise due to the underlying model architecture, typically convolutional neural networks, and are not a direct problem of the adversarial network \citep{2017arXiv170100160G}. Furthermore, generated image resolution has remained relatively low; however, recent improvements outlined in section \ref{sec:proggans} show improvements with regard to resolution.

\subsection{Evaluation}
The evaluation procedure for GANs remains an open debate as there exist no clear quantitative measurement to evaluate the models \citep{2017arXiv170100160G}. For example, models with high likelihood may produce unrealistic samples and models with low likelihood may produce realistic samples \citep{2017arXiv170100160G}. Several variations of the GANs examine the tradeoffs of divergence measures other than the original Kullback-Leibler divergence \citep{2015arXiv151101844T}. It has been shown that different metrics lead to different tradeoffs and different evaluations favor different models \citep{2015arXiv151101844T}. Therefore, it is important to asses one's model within the context of the whole application \citep{2015arXiv151101844T}.

\subsection{Discrete Data}
GANs assume a model implementation that is differentiable due to it's use of back propagation. Therefore, this restricts their ability to generate discrete data. Goodfellow proposes several solutions such as algorithm modifications, discretization of outputs or the use of concrete distributions or Gumbel-softmax \citep{2017arXiv170100160G}.

\section{Training Improvements}
A substantial amount of research has been conducted within the past three years with regards to improving GAN training procedures. A paper released by Salimans \textit{et al.} while at OpenAI outlines several techniques that can be applied to improve the GAN training process \citep{2016arXiv160603498S}. As mentioned in section \ref{sec:instability}, the optimal solution for GANs is the convergence to the Nash equilibrium of the minimax game; however, gradient descent only attempts to minimize the cost function and not to find the Nash equilibrium. The following techniques introduced in the paper address the instability of GANs and attempt to increase the probability of convergence.

\subsection{Feature Matching}
Feature matching involves the definition of a new cost function for GANs so that the generator does not overtrain on the current discriminator \citep{2016arXiv160603498S}. Instead of optimizing on the output of the discriminator, the new objective requires the generator to minimize the difference between the expected value of some intermediate layer. The objective function can be described as $\Big|\Big|\mathbb{E}_{x\sim p_{data}(x)}\Big[f(x)\Big] - \mathbb{E}_{z\sim p_{z}(z)}\Big[f(G(z))\Big]\Big|\Big|_2^2$ where $f(x)$ is the output of the activation function of some layer of the discriminator. Experimental evidence gathered has demonstrated the usefulness of this technique, especially when the training process becomes instable \citep{2016arXiv160603498S}. Furthermore, feature matching is useful for semi-supervised GAN applications as will later be described in section \ref{sec:semisupervised}.

\subsection{Minibatch Discrimination} \label{sec:minidisc}
Minibatch discrimination attempts to minimize the issues caused by mode collapse. This problem occurs as the discriminator processes each value $x$ independently of each other value. The gradients of discriminator will all point in the same direction, thus leading to mode collapse. This technique involves the use of a minibatch in place of a single value to allow to discriminator to examine multiple data examples at one time. The discriminator is then able to examine the distance between each sample in conjunction with the samples themselves and look for abnormal minibatch distributions. Successful applications of this technique allow for visually appealing images to be generated much quicker than by previous training techniques \citep{2016arXiv160603498S}. Preliminary results also show great reduction of mode collapse issues \citep{2017arXiv170100160G}.

\subsection{Historical Averaging}
Historical averaging introduces an extra term for each player's cost function such that they are penalized when choosing new parameters that differ from the historical mean value \citep{2016arXiv160603498S}. The new term becomes $||\theta - \frac{1}{t}\sum_{i=1}^{t}\theta[i]||^2$ where $t$ is the total number of iterations performed and $\theta[i]$ is the value of the parameters at iteration $i$. This technique has been demonstrated to help low dimensionality games reach an equilibrium where normal gradient descent has failed \citep{2016arXiv160603498S}.

\subsection{Label Smoothing}
Label smoothing was a technique first introduced in the 1980s which involves replacing the 0 and 1 targets with \textit{smoother} values such as 0.1 and 0.9 respectively \citep{2016arXiv160603498S}. This technique is used to improve the networks resistance to adversarial examples. Specifically, it is recommended one use only one-sided label smoothing by using 0.9 label instead of 1 for the real data label. If one alters the label for fake data, problems may arise when $p_{data}(x)$ nears $0$ and $p_{model}(x)$ becomes large for some sample $x$ \citep{2016arXiv160603498S}.

\subsection{Virtual Batch Normalization}
Batch normalization is a technique that was first introduced in 2014 which involves applying normalization for each training minibatch \citep{2015arXiv150203167I}. It allows for much higher learning rates, a more stable training process and a reduction of problems associated with parameter initialization \citep{2015arXiv150203167I}. This normalization technique has proved useful for GAN training; however, it often causes a dependence relationship between the samples of the minibatch \citep{2016arXiv160603498S}. A new technique called virtual batch normalization (VBN) was developed where a reference batch of samples is randomly chosen at the start of the training process and used throughout to normalize each training sample. However, this technique is computationally expensive as it requires forwarding two minibatches and it is recommended to only use VBN for the generator \citep{2016arXiv160603498S}.

\section{Architecture Variations}
Research within the field of deep generative modeling has increased exponentially since the debut of GANs in 2014. Significant improvements are being made with respect to image quality and training stability. Hundreds of papers have been released since 2014 (see figure \ref{fig:CumulativeNumberofGANPapersbyMonth}) which propose numerous architectural variations.

\begin{figure}
\centering
\includegraphics[scale=0.43]{cumulative_gans}
\caption{The cumulative amount of papers written about generative adversarial networks by month. The graph shows the exponential growth of GANs within the past three years. Image maintained by Avinash Hindipur \citep{November2017GANZoo}.}
\label{fig:CumulativeNumberofGANPapersbyMonth}
\end{figure}

\subsection{Fully Connected GANs}
The first GAN paper by Goodfellow \textit{et al.} made use of relatively simple fully connected neural networks for the generator and discriminator models \citep{2017arXiv171007035C}. These are networks where each neuron in one layer is connected to every neuron in the following layer. This design is able to generate simple images and useful for proof of concept; however, several other architectures have overtaken this implementation \citep{2017arXiv171007035C}.

\subsection{Convolutional GANs}
Although convolutional neural networks (CNNs) seem well suited for the image generation tasks associated with GANs, difficulties associated with training initially constrained the use of convolutional models for both the discriminator and generator \citep{2017arXiv171007035C}. When implementing CNNs within a GAN, one is often unable to train the discriminator and generator to the same level of capacity and representational power when compared to supervised CNN models \citep{2017arXiv171007035C}. One solution proposed by Denton \textit{et al.} made use of Laplacian pyramids (LAPGAN) to address this problem \citep{2015arXiv150605751D}. A Laplacian pyramid is a type of multi-scale signal representation which repeatedly smooths and downsamples an image to form a sort of pyramid of images. At each layer of the pyramid, a conditional convolutional GAN (see section \ref{sec:condgan}) is trained given the generator output of the above layer.

Another solution to this problem was proposed by Radford \textit{et al.} made alterations to conventional deep convolutional network architecture \citep{2015arXiv151106434R}. Max pooling was replaced with strided layers within the discriminator and fractionally-strided layers within the generator. Furthermore, LeakyReLUs are used for all layers of the discriminator in place of the standard ReLU non-linearity function. The fractionally strided layers within the generator replicate the benefits of the progressive upscaling within the Laplacian pyramid \citep{2017arXiv171007035C}. The strided layers allow the GAN to learn both the down-sampling and up-sampling operators during training \citep{2017arXiv171007035C}. This architecture, called DCGAN, has become the base model for most GANs today \citep{2017arXiv170100160G}.

\subsection{Conditional GANs} \label{sec:condgan}
Initially proposed by Goodfellow as a straightforward extension of the first GAN architecture \citep{2014arXiv1406.2661G}, Mirza \textit{et al.} successfully implemented a conditional generative model in late 2014 \citep{2014arXiv1411.1784M}. A generative model differs from a conventional models as an additional parameter $c$ is given to both the generator $G(z|c)$ and discriminator $D(x|c)$ to condition the GAN. The new objective function of a conditional GAN becomes:

@@

These architectures are better suited for multi-modal output as they are able to provide better data representations \citep{2017arXiv171007035C}. By conditioning the GAN, it is possible to control the data generation process. This additional input may be class label or other items such as text or an image. In section \ref{sec:img2img}, we will see how conditional GANs are applied to image translation tasks.

>>>>

\subsection{Image Quality}
When generating images, problems often arise due to the generators difficulties with perspective, counting and structure \citep{2017arXiv170100160G}. For example, an generated animal may  have a 2D structure, misplaced limbs or too many of a specific body part. These issues arise due to the underlying generator architecture, typically deconvolutional neural networks, and are not an inherent problem of the adversarial network \citep{2017arXiv170100160G}. Furthermore, generated image resolution has remained relatively low; however, recent improvements outlined in section \ref{sec:proggans} show drastic improvements with regard to resolution.

\subsection{Evaluation}
The evaluation procedure for GANs remains an open debate as there exist no clear quantitative measurement to evaluate generative models \citep{2017arXiv170100160G}. For example, models with a high likelihood may produce unrealistic samples whereas models with a low likelihood may produce realistic samples \citep{2017arXiv170100160G}. It has been shown that different metrics lead to different tradeoffs and different evaluations favor different models \citep{2015arXiv151101844T}. Therefore, it is important to asses one's generative model within the context of the whole application \citep{2015arXiv151101844T}.

\subsection{Discrete Data}
GANs assume a model implementation that is differentiable due to their use of back propagation. Therefore, this restricts their ability to generate discrete data. Goodfellow proposes several solutions such as algorithm modifications, discretization of outputs or the use of concrete distributions or Gumbel-softmax \citep{2017arXiv170100160G}.

\section{Training Improvements}
A substantial amount of research has been conducted within the past three years with regards to improving GAN training procedures. A paper released by Salimans \textit{et al.} while at OpenAI outlines several techniques that can be applied to improve the GAN training process \citep{2016arXiv160603498S}. As mentioned in section \ref{sec:instability}, the optimal solution for GANs is the convergence to the Nash equilibrium of the minimax game; however, gradient descent attempts to minimize the cost function and will not necessarily find the Nash equilibrium. The following techniques introduced in the paper address the instability of GANs and attempt to increase the probability of convergence.

\subsection{Feature Matching}
Feature matching involves the definition of a new cost function for GANs to reduce the probability that the generator will not overtrain on the current discriminator \citep{2016arXiv160603498S}. Instead of optimizing on the output of the discriminator, the new objective requires the generator to minimize the difference between the expected value of some intermediate layer. The new objective function can be described as $\Big|\Big|\mathbb{E}_{x\sim p_{data}(x)}\Big[f(x)\Big] - \mathbb{E}_{z\sim p_{z}(z)}\Big[f(G(z))\Big]\Big|\Big|_2^2$ where $f(x)$ is the output of the activation function of some layer of the discriminator. Experimental evidence gathered has demonstrated the usefulness of this technique, especially when the training process becomes instable \citep{2016arXiv160603498S}. Furthermore, feature matching is useful for semi-supervised GAN applications as will later be described in section \ref{sec:semisupervised}.

\subsection{Minibatch Discrimination} \label{sec:minidisc}
Minibatch discrimination attempts to minimize the issues caused by mode collapse. This issue occurs as the discriminator processes each sample $x$ independently of each other sample. Therefore, the gradients of discriminator will all point in the same direction, thus leading to mode collapse. This technique involves the use of a minibatch in place of a single sample to allow to discriminator to examine multiple data examples at one time. The discriminator is then able to examine the distance between each sample in conjunction with the samples themselves and look for abnormal minibatch distributions. Successful applications of this technique allow for visually appealing images to be generated much quicker than previous training techniques \citep{2016arXiv160603498S}. Preliminary results also show a great reduction of mode collapse \citep{2017arXiv170100160G}.

\subsection{Historical Averaging}
Historical averaging introduces an extra term to each player's cost function such that they are penalized when choosing new parameters that differ from the historical mean value \citep{2016arXiv160603498S}. The new term added is $||\theta - \frac{1}{t}\sum_{i=1}^{t}\theta[i]||^2$ where $t$ is the total number of iterations performed and $\theta[i]$ is the value of the parameters at iteration $i$. This technique has been demonstrated to help low dimensionality games reach the Nash equilibrium where normal gradient descent would have failed \citep{2016arXiv160603498S}.

\subsection{Label Smoothing}
Label smoothing is a technique that was first introduced in the 1980s which involves replacing the 0 and 1 targets with \textit{smoother} values such as 0.1 and 0.9 respectively \citep{2016arXiv160603498S}. This technique is used to improve the networks resistance to adversarial examples. Specifically, it is recommended to use only one-sided label smoothing by replacing the 0.9 label instead of 1 for the real data label. If one alters the label for fake data, problems may arise when $p_{data}(x)$ nears $0$ and $p_{model}(x)$ becomes large for some sample $x$ \citep{2016arXiv160603498S}.

\subsection{Virtual Batch Normalization}
Batch normalization (BN), a technique that was first introduced in 2014, involves normalizing each training minibatch \citep{2015arXiv150203167I}. It allows for much higher learning rates, a more stable training process and a reduction of problems associated with parameter initialization \citep{2015arXiv150203167I}. BN has proved useful for GAN training; however, it often causes a dependence relationship between the samples of the minibatch \citep{2016arXiv160603498S}. A new technique called virtual batch normalization (VBN) resembles BN; however, a reference batch of samples is randomly chosen at the start of the training process and used throughout to normalize each training sample. This technique is computationally expensive as it requires forwarding two minibatches and it is recommended that one only use VBN for the generator \citep{2016arXiv160603498S}.

\section{Architecture & Training Variations}
Research within the field of deep generative modeling has increased exponentially since the debut of GANs in 2014. Significant improvements have been made with respect to image quality and training stability. Hundreds of papers have been released on the subject of GANs (see figure \ref{fig:CumulativeNumberofGANPapersbyMonth}) which propose numerous architectural and training variations.

\begin{figure}
\centering
\includegraphics[scale=0.43]{cumulative_gans}
\caption{The cumulative amount of papers written about generative adversarial networks (GANs) by month. The graph shows the exponential growth of GAN research that has occurred within the past three and a half years. Image maintained by Avinash Hindipur \citep{November2017GANZoo}.}
\label{fig:CumulativeNumberofGANPapersbyMonth}
\end{figure}

\subsection{Fully Connected GANs}
The first GAN paper by Goodfellow \textit{et al.} made use of relatively simple fully connected neural networks for the generator and discriminator models \citep{2017arXiv171007035C}. These are networks where each neuron in one layer is connected to every neuron in the following layer. This design is able to generate simple images and is useful for proof of concept; however, several other architectures regularly outperform this implementation \citep{2017arXiv171007035C}.

\subsection{Convolutional GANs}
Although convolutional neural networks (CNNs) seem well suited for the image generation tasks often associated with GANs, difficulties with training initially constrained the use of convolutional models for both the discriminator and generator \citep{2017arXiv171007035C}. When implementing CNNs within a GAN, one is unable to train the discriminator and generator to the same level of capacity and representational power as supervised CNN models \citep{2017arXiv171007035C}. One solution proposed by Denton \textit{et al.} made use of Laplacian pyramids to address this problem \citep{2015arXiv150605751D}. A Laplacian pyramid is a type of multi-scale signal representation which repeatedly smooths and downsamples an image to form a sort of pyramid of images. At each layer of the pyramid, a conditional convolutional GAN (see section \ref{sec:condgan}) is trained given the generator output of the above layer.

Another solution to this problem proposed by Radford \textit{et al.} made alterations to the conventional deep convolutional network architecture \citep{2015arXiv151106434R}. Max pooling was replaced with strided layers within the discriminator and fractionally-strided layers within the generator. Furthermore, LeakyReLUs were used for all layers of the discriminator in place of the standard ReLU non-linearity function. The fractionally strided layers within the generator replicate the benefits of the progressive upscaling within the Laplacian pyramid \citep{2017arXiv171007035C}. The strided layers allow the GAN to learn both the down-sampling and up-sampling operators during training \citep{2017arXiv171007035C}. This architecture, called DCGAN, has become the base model for most GANs today \citep{2017arXiv170100160G}.

\subsection{Conditional GANs} \label{sec:condgan}
Initially proposed by Goodfellow as a straightforward extension of the first GAN architecture \citep{2014arXiv1406.2661G}, Mirza \textit{et al.} successfully implemented a conditional GAN (CGAN) in late 2014 \citep{2014arXiv1411.1784M}. CGANs differ from conventional models due to the additional parameter $c$ that is given to both the generator $G(z|c)$ and discriminator $D(x|c)$ to condition the GAN. The new objective function of a CGAN becomes:

@@

These architectures are better suited for multi-modal output (see section \ref{sec:img2img}) as they are able to provide better data representations \citep{2017arXiv171007035C}. By conditioning the GAN, it is possible to control the data generation process. This additional input may be class label or other items such as text or an image. For example, given that a CGAN was conditioned on sentences, one may be able to describe a scenario and have that scenario synthesized as an image \citep{2016arXiv160505396R}. In section \ref{sec:img2img}, we will see how conditional GANs are applied to image translation tasks.
